---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# wcvpdata

<!-- badges: start -->
[![R-CMD-check](https://github.com/PaulESantos/wcvpdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PaulESantos/wcvpdata/actions/workflows/R-CMD-check.yaml)
[![R-universe](https://PaulESantos.r-universe.dev/badges/wcvpdata)](https://PaulESantos.r-universe.dev/wcvpdata)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

## Overview

`wcvpdata` provides a local R interface to the **World Checklist of
Vascular Plants (WCVP)**, a comprehensive global database of accepted
plant names and their synonyms maintained by the Royal Botanic Gardens,
Kew. This package bundles standardized snapshots of the WCVP taxonomic
and geographic distribution data (Version 16, June 2026), enabling
large-scale botanical research and biodiversity analysis directly in R
— without requiring repeated downloads or internet access.

<invoke name="25-1">The WCVP database was initiated in 1988 by Rafaël Govaerts and has grown into an **international collaborative programme** with editors,
compilers, and reviewers from around the globe.

## Installation

You can install the latest development version of wcvpdata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("PaulESantos/wcvpdata")

# From r-universe
install.packages("wcvpdata",
  repos = c(
    "https://paulesantos.r-universe.dev",
    "https://cloud.r-project.org"
  )
)
```
## Usage

Load the library:
```{r}
library(wcvpdata)
```

### Checking the Database Version
To see what version of the WCVP database is bundled with the package:
```{r}
wcvp_version()
```

You can check if the bundled version matches the latest one uploaded to the Kew SFTP server:
```{r, eval = TRUE}
wcvp_check_version()
```

### Validating Local Data Integrity
To verify that the locally installed datasets are complete and match the expected package metadata dimensions:
```{r}
wcvp_validate_data()
```


## Citation

If you use wcvpdata, please cite the snapshot of the database you used and the paper describing WCVP.

You can find out how to cite these by running:
```{r}
print(citation("wcvpdata"), bibtex=TRUE)
```


